---
title: "Turkey's Visitor Behavior"
author: 
  - Aykal, Aslı Nil[^1]
bibliography: ../bibliography/biblio.bib
csl: ../csl/apa.csl
header-includes:
  - \usepackage{polyglossia}
  - \setmainlanguage{english}
  - \usepackage{booktabs}
  - \usepackage{caption} 
  - \captionsetup[table]{skip=10pt}
output:
  bookdown::pdf_document2:
    fig_caption: yes
    fig_height: 3
    fig_width: 4
    keep_tex: no
    latex_engine: xelatex
    number_sections: yes
    toc: no
geometry: margin=1in
link-citations: yes
urlcolor: blue
fontsize: 12pt
biblio-style: apalike
abstract: |
 With its geographical, cultural, historical and touristic beauties, Turkey receives visitors from various parts of the world every year for many reasons. There is a noticeable increase and decrease in the number of these visitors in some periods. The main purpose of this study is to investigate the factors affecting the visitor demand to Turkey in the last five years. In this project, various literature studies were used to understand the factors on which visitor demand depends, and the data were obtained from the website of the Central Bank. In the research, both qualitative and quantitative research methods were used and the effects of factors such as inflation, Covid-19 and terrorist incidents on the visitor demand to Turkey were examined. As a result of this research, a decrease was observed in the number of visitors during the Covid-19 period, it was determined that inflation did not have a definite effect on the visitor demand, and it was observed that the terrorist events that occurred had a negative impact on the visitor demand.
 
  KEY WORDS:
  Tourism, Tourism in Turkey, Tourism demand, Turkey's tourism demand, Covid-19,The relationship between inflation and tourism demand





---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE, message = FALSE, warning = FALSE)
```

<!-- ======================================================================= -->
<!-- ============================== FOOTNOTES ============================== -->
<!-- ======================================================================= -->
[^1]: 20080470, [Github Repo](https://github.com/aslaykal/final.git)




# Introduction
Turkey has been a country that has been attracting tourists for years with its cultural and natural beauties. With its culture, geopolitical location, climate, natural and artificial heritage, history, gastronomy and diverse social structure, it attracts attention from a large part of the world, and in this sense, it positively affects tourism statistics. Compared to many countries, Turkey has higher visitor statistics for the reasons mentioned. However, in recent years, fluctuations rather than stability have been observed in the number of visitors to Turkey. In this study, it has been investigated how the number of foreign visitors in Turkey has changed in the last five years and what factors contributed to this change. For this analysis, a data set consisting of 62 observers and 10 variables obtained by the Central Bank of the Republic of Turkey was used. By using the monthly values of the last 5 years, the effects of factors such as the covid-19 process, inflation rate, terrorist acts, health tourism and common religious identity that shape tourism statistics and how they affect them were investigated.




## Literature Review
Recently, certain fluctuations have been observed in Turkey's visitor statistics, especially in certain periods. Factors such as the pandemic process that came with the Covid-19 epidemic, the increase in the inflation rate, the terrorist acts in recent years, the change in the demand for health tourism recently, and the common culture-belief have a direct or indirect effect on these fluctuations.

Pektaş and Ünlüönen (2019), analyzed the relationship between inflation and tourism performance in Turkey and observed that tourism demand increased between 2017-2018 when Turkey had the highest inflation level at the time of the research. In the period of increased inflation in Turkey between 2004 and 2018, tourism revenues and the number of foreign tourists also increased. In addition, according to the CPI data analyzed between 2004-2018, inflation is not a direct determinant of the total number of tourists and tourism revenues. However, they also stated that inflation cannot be the only factor affecting tourism performance.
[@pektas:2020]

Çeti and Ünlüönen (2019), on the other hand, made an economic evaluation of health tourism by taking advantage of the number of tourists coming to Turkey to receive health services and taking into account the expenditures of these tourists. As a result of their findings, an increase has been observed in the number of visitors coming to Turkey for health purposes in recent years. In 2003, the ratio of the number of tourists coming for health purposes to the total number of tourists was determined as 0.86%, and this percentage increased to 1.21% at the end of 2018. 2018 was the year in which the ratio of the number of tourists coming for health purposes to the total number of tourists was the highest. As a result, the demand for health tourism has increased for Türkiye.[@buse2020economic]

Bilgiçli and Çolaksel (2019) indicates that the reason why Turkey is especially popular with visitors from the Middle East is that it creates a positive image as a country of religion, cultural and historical ties, shopping-fashion, TV series, health and vacation. Proximity to Arabs in terms of belief and culture, Islamophobic perspectives of Europe and America, especially after September 11, seeing Turkey as a modern Muslim country, affordable shopping opportunities and prices, seeing Turkey as a role model, Arab investors' interest in Turkey. The reasons such as the increase in the number of Arabs buying houses with the interest in Turkey's culture, historical beauties and nature are also important factors in Turkey's receiving a large number of Middle Eastern visitors.[@bilgiccli2019orta]

Bassil, Saleh, and Anwar (2017), examine the effects of local and transnational terrorism on the number of visitors to Lebanon, Turkey and Israel. This factor has a negative impact on visitor numbers, as terrorism makes tourists nervous about personal safety and security issues. In addition, the empirical results analyzed between 1995 and 2007 showed that terrorism in a country caused a decrease in the number of visitors not only to that country but also to other countries. Especially the terrorist incidents in Israel caused changes in the number of tourists coming to Turkey and Lebanon. In their findings, they stated the total effect of local and transnational terrorism in Turkey on the logarithm of Turkey's visitor numbers as -0.057 and -0.135, respectively. This revealed a decrease of 18% in the total number of visitors.[@bassil2019terrorism]

Günay, Bayraktaroğlu, Özkul (2020), The pandemic process that came with the Covid-19 epidemic caused countries to impose international mobility restrictions as a precaution against the virus. Taking these measures to stop the spread of the virus has almost stopped international mobility and tourism activities have been postponed. However, Özdemir and Yıldız (2020) state that the tourism sector is the most affected area as a result of the adverse conditions of the Pandemic, and that the COVID-19 outbreak has changed the future travel and vacation plans and the understanding of tourists.[@gunay2020assessing;@ozdemir2020covid]



# Data
The data to be used in this analysis are obtained from the Central Bank data, these data cover the last 5 years, consist of 62 observers and 10 variables. For the research, the data of the last 5 years will be examined monthly. Variables include Albania, Holland, Russia, Brazil, Canada, South Africa, Egypt, China, Saudi Arabia and Australia. These variables were chosen to cover almost every continent around the world and their relations with tourism demand were examined based on factors such as inflation, Covid-19 and terrorist events. A statistical summary of the data is given in Table 1:


```{r include=FALSE}
albania <- c(6875, 5070, 8203, 7856, 7664, 11992, 15150, 20076, 15642, 9780, 8539, 9616, 7358, 6319, 10607, 9531, 9270, 11930, 15264, 20556, 16141, 10549, 8654, 9518, 8125, 5903, 3878, 534, 572, 950, 3680, 6177, 5368, 6184, 5517, 4468, 3135, 3183, 4358, 4213, 4076, 8233, 11395, 15590, 10471, 8424, 8906, 7060, 6508, 6107, 8789, 8601, 9918, 13744, 19984, 24172, 18560, 11094, 10744, 7384, 7602)

holland <- c(29740, 22900, 30784, 50221, 138584, 77747, 94425, 252882, 123659, 131238, 37731, 30558, 32356, 27827, 36737, 69311, 139185, 96915, 107608, 272857, 129384, 142721, 39871, 31420, 33980, 33439, 21832, 447, 244, 4301, 16482, 75010, 29506, 26984, 18967, 17001, 12643, 10426, 13876, 15748, 17445, 23106, 42186, 219504, 91142, 105698, 44838, 36079, 54290, 32346, 49496, 50781, 167769, 96805, 110168, 325936, 143845, 150041, 46385, 34446, 43193)

russia <- c(80728, 62970, 85145, 169550, 714441, 972985, 947371, 914725, 912817, 788144, 207158, 81816, 114310, 80827, 123671, 232488, 856511, 1021201, 1055387, 1078045, 1057296, 980589, 288156, 111926, 154941, 101284, 94333, 1413, 1512, 3083, 8051, 294313, 601864, 590914, 198463, 74202, 110237, 76382, 182245, 204845, 20684, 42060, 703659, 898620, 935117, 1031717, 343576, 109617, 167451, 101964, 98634, 99576, 288225, 535404, 716435, 805145, 794483, 777034, 429728, 266435, 313267)

canada <- c(6368, 5039, 6678, 7534, 8323, 8910, 12183, 15410, 13620, 10695, 7571, 6858, 7153, 6057, 8266, 10126, 11125, 12573, 17102, 21356, 16653, 14437, 9640, 7482, 7680, 6510, 4252, 317, 40, 216, 1470, 2621, 3819, 3396, 2993, 2872, 2383, 1742, 1533, 2052, 2198, 3438, 7216, 12588, 11662, 10514, 8845, 7577, 6854, 5294, 9129, 8929, 16813, 18225, 24013, 30184, 27570, 28076, 16488, 10466, 10910)

south_africa <- c(3184, 1775, 2430, 4442, 4042, 4507, 7466, 4522, 6523, 5955, 2655, 4092, 4011, 1923, 4278, 5411, 5000, 9861, 10794, 6717, 9840, 8386, 3410, 4313, 5134, 2755, 1914, 3, 65, 21, 207, 610, 707, 941, 941, 872, 348, 194, 220, 311, 249, 384, 759, 745, 633, 809, 774, 614, 940, 2184, 4662, 3762, 5283, 6527, 10413, 7008, 8443, 10772, 4868, 4443, 5605)

egypt <- c(9632, 8922, 9105, 11239, 9432, 9953, 13155, 15679, 15116, 12371, 10676, 10973, 13249, 11679, 12159, 14509, 10070, 14492, 17182, 20288, 16325, 14455, 12067, 12544, 14311, 11307, 5704, 82, 254, 602, 1389, 3935, 5450, 8891, 9059, 6368, 5165, 4757, 5518, 6087, 3632, 6977, 12679, 13198, 15033, 15007, 12726, 13226, 13257, 15439, 17431, 7553, 16192, 19181, 25830, 23894, 27283, 23869, 17536, 16155, 15320)

china <- c(25722, 32714, 29029, 34355, 35844, 34327, 31944, 40058, 34458, 44831, 32532, 24852, 23968, 33839, 28897, 37162, 43057, 43997, 41967, 46983, 38458, 49297, 28782, 23431, 24306, 8962, 2118, 57, 46, 64, 577, 1242, 1621, 1619, 1110, 2449, 2162, 1457, 2470, 1559, 1903, 2301, 2675, 3459, 3312, 3416, 3528, 3758, 3338, 3488, 3115, 6785, 7614, 9313, 10820, 10081, 9042, 8939, 8328, 10216, 10608)
saudi_arabia <- c(51511, 20375, 32668, 48575, 42035, 37098, 141491, 179420, 76502, 44961, 27056, 23812, 46006, 19368, 27901, 39878, 19718, 54832, 121563, 149821, 42209, 19806, 20533, 17975, 53927, 15482, 6173, 250, 468, 89, 69, 350, 688, 694, 529, 328, 391, 191, 223, 388, 318, 833, 1908, 2115, 795, 924, 807, 933, 829, 964, 1490, 543, 2145, 4191, 56355, 170780, 85099, 60864, 42860, 62545, 34385)

australia <- c(4396, 2964, 3047, 5670, 6333, 8693, 14288, 13527, 14748, 12028, 5770, 5026, 5809, 3413, 3688, 7836, 9173, 12022, 17737, 16243, 18230, 14653, 7551, 5437, 6265, 3819, 2666, 43, 118, 218, 291, 621, 1011, 956, 622, 649, 430, 403, 503, 673, 615, 636, 1234, 1383, 1276, 2338, 1954, 2698, 3799, 2866, 3590, 4267, 6914, 11050, 15992, 18324, 18174, 17142, 11135, 6214, 6402)

brazil <- c(4081, 3398, 3306, 5834, 8543, 8289, 7921, 6901, 9047, 8514, 6686, 5860, 6822, 5033, 7209, 6546, 9768, 9788, 10449, 8711, 12175, 11521, 6914, 5748, 7898, 4779, 2837, 25, 126,64, 426, 836, 1314, 1806, 1954, 1611, 1672, 677, 540, 573, 719, 825, 1328, 1734, 1728, 2323, 2206, 3329, 3673, 2885, 3945, 5199, 7730, 8451, 9073, 7647, 11358, 11940, 7468, 4501, 5182)



Date <- c("2018-01", "2018-02", "2018-03", "2018-04", "2018-05", "2018-06", "2018-07", "2018-08", "2018-09", "2018-10", "2018-11", "2018-12", "2019-01", "2019-02", "2019-03", "2019-04", "2019-05", "2019-06", "2019-07", "2019-08", "2019-09", "2019-10", "2019-11", "2019-12", "2020-01", "2020-02", "2020-03", "2020-04", "2020-05", "2020-06", "2020-07", "2020-08", "2020-09", "2020-10", "2020-11", "2020-12", "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12", "2022-01", "2022-02", "2022-03", "2022-04", "2022-05", "2022-06", "2022-07", "2022-08", "2022-09", "2022-10", "2022-11", "2022-12","2023-01")

Date <- as.Date(paste0(Date, "-01"), format = "%Y-%m-%d")

data <- data.frame(Date = Date, albania = albania, holland = holland, russia = russia, brazil = brazil, canada = canada, south_africa =south_africa, egypt = egypt, china = china, saudi_arabia = saudi_arabia, australia = australia)


```

```{r include=FALSE}
data_without_date <- data[, !colnames(data) %in% "Date"]
summary_table <- data.frame(
  Mean = sapply(data_without_date, mean),
  StandardDeviation = sapply(data_without_date, sd),
  Minimum = sapply(data_without_date, min),
  Maximum = sapply(data_without_date, max)
)

print(summary_table)
```


```{r echo=FALSE}
library(kableExtra)
kable(summary_table, caption = "Summary Statistics", format = "markdown")
```
\newpage
# Methods and Data Analysis
## Data Analysis
In Figure 1,2 and 3, According to the data received from the Central Bank, the number of visitors from Holland, Egypt and Russia to Turkey from September 2022 to the end of January 2023 is shown. Accordingly, an increase in the number of visitors in Holland from September to October, a serious decrease from October to November and a decrease from November to December were observed. Then, a small increase was observed in January 2023. In Egypt's data, a regular decrease was observed from September to January, and the most striking decrease was experienced in November. On the other hand, there were decreases in Russian data from September to December, a small increase was observed in January, and the most striking decrease was reflected in November.


```{r echo=FALSE, fig.height=3, fig.width=3}
library(ggplot2)

subset_data <- tail(data, 5)

ggplot(subset_data, aes(x = Date, y = holland)) +
  geom_col(fill = "blue") +
  labs(x = "Date", y = "number of tourists", title = "Figure 1: Visitors From the Holland") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1) , text = element_text(size = 8))





library(ggplot2)

ggplot(subset_data, aes(x = Date, y = russia)) +
  geom_col(fill = "red") +
  labs(x = "Date", y = "number of tourists", title = "Figure 2: Visitors From the Russia") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1) , text = element_text(size = 8))
```

```{r echo=FALSE, fig.height=3, fig.width=3}
library(ggplot2)
ggplot(subset_data, aes(x = Date, y = egypt)) +
  geom_col(fill = "violet") +
  labs(x = "Date", y = "number of tourists", title = "Figure 3: Visitors From the Egypt") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1) , text = element_text(size = 8))
```

In Figure 4, According to the data received from the Central Bank, the data of six countries, namely Albania, Canada, China, Holland, Russia and Saudi Arabia, were examined and the changes in the tourism demands of these countries were examined. In the period from January 2020 to October 2020, stable decreases were experienced in all six countries from January to April, and tourism demands of all six countries reached the lowest level in 5 years in April 2020. After July, especially Russia's tourism demand to Turkey increased dramatically and an increase was observed in other countries as well.


```{r echo=FALSE, fig.height=6, fig.width=8}
library(ggplot2)
filtered_data <- subset(data, Date >= as.Date("2019-12-01") & Date <= as.Date("2020-12-1"))

ggplot(filtered_data, aes(x = Date)) +
  geom_line(aes(y = albania, color = "albania")) +
  geom_line(aes(y = china, color = "china")) +
  geom_line(aes(y = canada, color = "canada")) +
  geom_line(aes(y = russia, color = "russia")) +
  geom_line(aes(y = holland, color = "holland")) +
  geom_line(aes(y =  saudi_arabia, color = "saudi arabia " )) +                     labs(x = "Date", y = "countries", color = "Category") +
  scale_color_manual(values = c("black", "blue", "green", "orange", "purple","yellow"))

```

\newpage


In Figure 5 and 6, According to the data received from the Central Bank, Turkey's total visitor demand and inflation rates in the last 5 years are given based on ten countries. Although increases and decreases were observed in the inflation rate between 2018-2019, there was a general increase and an increase was observed in the number of visitors, especially in the summer months. Although the inflation rate was at its highest level in January 2019, the number of visitors increased compared to the previous month in the corresponding period.In the 2020-2021 period, fluctuations in the inflation rate were observed and the number of visitors increased again in the summer months, but a decrease was observed in this increase compared to previous years. There were fluctuations in inflation between 2021-2022, and the number of visitors increased again in the summer months, and this increase is higher than the previous period. Between 2022 and 2023, inflation reached its highest level in October 2022, although a decrease was observed in the number of visitors in the same month. Again, while inflation fluctuated between these years, an increase was observed in the number of visitors during the summer months.

```{r echo=FALSE}
 Date <- c("2018-01", "2018-02", "2018-03", "2018-04", "2018-05", "2018-06", "2018-07", "2018-08", "2018-09", "2018-10", "2018-11", "2018-12", "2019-01", "2019-02", "2019-03", "2019-04", "2019-05", "2019-06", "2019-07", "2019-08", "2019-09", "2019-10", "2019-11", "2019-12", "2020-01", "2020-02", "2020-03", "2020-04", "2020-05", "2020-06", "2020-07", "2020-08", "2020-09", "2020-10", "2020-11", "2020-12", "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12", "2022-01", "2022-02", "2022-03", "2022-04", "2022-05", "2022-06", "2022-07", "2022-08", "2022-09", "2022-10", "2022-11", "2022-12","2023-01")

Date <- as.Date(paste0(Date, "-01"), format = "%Y-%m-%d")
  aylikyüzde <- c(6.65, 1.18, 2.88, 3.54, 3.08, 1.46, 2.37, 4.95, 2.98, 7.25, 5.46, 4.81, 11.10, 13.58, 3.51, 2.39, 1.25, 1.12, 1.80, 1.94, 0.89, 1.68, 1.08, 0.91, 1.68, 1.25, 2.30, 2.13, 0.97, 0.86, 0.58, 1.13, 1.36, 0.85, 0.57, 0.35, 1.35, 0.74, 0.38, 2, 0.99, 0.86, 1.36, 0.03, 0.95, 1.69, 1.03, 0.16, 1.06, -0.40, -1.44, 2.67, 6.30, 2.30, 0.55, 2.61, 1.62, 1.87, 0.99, 0.73, 1.02)
enflasyon <- data.frame(Date = Date, aylikyüzde = aylikyüzde)
library(ggplot2)
plot(Date, aylikyüzde, type = "l", xlab = "Date", ylab = "Visitor Count", 
     main = "Figure 5:Inflation Rate", xlim = c(min(Date), max(Date)))




```

```{r echo=FALSE}
Dates <- c( "2018-01", "2018-02", "2018-03", "2018-04", "2018-05", "2018-06", "2018-07", "2018-08", "2018-09", "2018-10", "2018-11", "2018-12", "2019-01", "2019-02", "2019-03", "2019-04", "2019-05", "2019-06", "2019-07", "2019-08", "2019-09", "2019-10", "2019-11", "2019-12", "2020-01", "2020-02", "2020-03", "2020-04", "2020-05", "2020-06", "2020-07", "2020-08", "2020-09", "2020-10", "2020-11", "2020-12", "2021-01", "2021-02", "2021-03", "2021-04", "2021-05", "2021-06", "2021-07", "2021-08", "2021-09", "2021-10", "2021-11", "2021-12", "2022-01", "2022-02", "2022-03", "2022-04", "2022-05", "2022-06", "2022-07", "2022-08", "2022-09", "2022-10", "2022-11", "2022-12", "2023-01")
Thenumberofvisitors <- c(1601862, 1444008, 1862961, 2433008, 3398215, 4023879, 4857186, 6001678, 5045870, 4292466, 2234062, 1756707, 1804135, 1574614, 2081631, 2845084, 3832240, 4821887, 5717804, 7018257, 5702078, 4878255, 2501695, 1935290, 2065570, 1664208, 906067, 50630, 51588, 128861, 529990, 1712380, 2187683, 2000557, 960402, 681408, 570447, 494337, 761443, 895342, 757518, 1491955, 3231555, 4578265, 3648131, 3777732, 2080510, 1652786, 1593112, 1400014, 1950838, 2381722, 3483143, 4517203, 5693111, 7076690, 5856202, 5206943, 2937397, 2245147, 2323115)
visitors <- data.frame(Dates = Dates, Thenumberofvisitors = Thenumberofvisitors)
Dates <- as.Date(paste0(Dates, "-01"), format = "%Y-%m-%d")
Thenumberofvisitors <- as.numeric(Thenumberofvisitors)
plot(Dates,Thenumberofvisitors, type = "l", xlab = "Date", ylab = "Visitor Count", 
     main = "Figure 6 :Number of Visitors ", xlim = c(min(Dates), max(Dates)))
```

## Methods

Correlation analysis was performed to understand the relationships of the variables with each other. Accordingly, the following result was obtained:

Correlation coefficients of each country with other countries are represented by the values in the data set. For example, the correlation coefficient between "Albania" and "Holland" was found to be 0.8336122. Correlation coefficients between other countries are also given in Figure 7. Blue majority results indicate high positive correlations. In other words, there is a strong relationship between the countries concerned.
```{r echo=FALSE}
library(corrplot)
cor_matrix <- cor(data[,-1])
corrplot(cor_matrix, method = "circle", tl.col = "black", tl.cex = 0.7)
```

# Conclusion
In this study, the effect of the factors that cause changes in the number of visitors in Turkey was investigated. In this context, various literature studies have been used to understand the factors affecting Turkey's visitor demand, the data has been taken from the website of the Central Bank and various graphs have been prepared with these data, making it easier to understand and observe the data. Various analyzes were made as a result of the quantitative and qualitative methods used in the study. The important results of these analyzes are as follows:

After the terrorist act in November 2022, a serious decrease was observed in the tourism demand of all three countries in the period from October to December. Accordingly, there is a negative relationship between terrorist acts and visitor demand.


Due to the Covid-19 pandemic process, a serious decrease was observed in the tourism demands of all six countries in the period from January 2020 to April 2020.

There is no clear relationship between inflation and visitor demand. Although the majority of the data explains that there is a positive relationship between inflation and tourism demand, the level of inflation increased in October 2022, but a decrease was observed in the number of visitors in the same month. Therefore, it is concluded that inflation is not a direct factor on visitor demand.

Using the data obtained from the correlation test, we cannot determine exactly how the number of foreign visitors in Turkey has changed and which factors contributed to this change, because only the correlations between the countries are given in the table. These relationships do not give the factors affecting the change in visitor numbers. More factors need to be considered for this.

In this research, how the number of foreign visitors in Turkey has changed in the last five years and what factors contributed to this change were examined. Three factors, namely the inflation rate, the Covid-19 pandemic process and terrorist acts, shaped the data study. Sufficient data on factors such as common religious identity and health tourism could not be reached and examined. In order to improve this study, more research can be done on these factors and the scope of the research can be expanded. In addition, it should be noted that the results cannot be generalized since this study covers 10 randomly selected countries around the world. In order to reach a more comprehensive result worldwide, studies with more data can be carried out and micro-studies can be made to support these studies.
\newpage

# References {#references}
<div id="refs"></div>

